Hypertension and microbiome

### 2018-12-31 rrbiome.Rmd
Better subgrouping real

filter.phenotype.data <- function(pseq,
                                  included = c("DNA_OK", "MEN", "BL_AGE", "SYSTM", "DIASM", "BP_TREAT", "BMI",
                                               "CURR_SMOKE", "DIAB", "BL_USE_RX_C09", "Q57X", "BL_USE_RX_C03",
                                               "BL_USE_RX_C07", "BL_USE_RX_C08"))
    meta(pseq) %>%
        tibble::rownames_to_column(var = "rowname") %>%
        dplyr::select(rowname, included) %>%
        dplyr::mutate(PULSEPRESSURE = SYSTM - DIASM,
                      HYPERTENSION = factor(ifelse(SYSTM >= 140 | DIASM >= 90 | BP_TREAT == 1, 1, 0)),
                      SEX = factor(ifelse(MEN == "Female", 1, 0)),
                      MAP = 2./3.*DIASM + 1./3.*SYSTM) %>%
 #       dplyr::filter(DNA_OK == 1) %>%
        stats::na.omit() %>%
        dplyr::select(-MEN, -DNA_OK) %>%
        tibble::remove_rownames() %>%
        tibble::column_to_rownames(var = "rowname")
```

@@

Number of participants in each combination of categorical variables

```{r numbers by group}
dset.species %>%
    group_by(SEX, BP_TREAT, CURR_SMOKE,  HYPERTENSION, Q57X,
             BL_USE_RX_C09, BL_USE_RX_C03, BL_USE_RX_C07, BL_USE_RX_C08) %>%
    select(-SYSTM, -DIASM, -BL_AGE, -BMI, -PULSEPRESSURE, -MAP) %>%
    summarise(n = n()) %>%
    arrange(desc(n)) %>%
    kable            
```

Number of particiapants in each combination of subset of categorical variables

```{r numbers by group reduced}
dset.species %>% group_by(SEX, BP_TREAT, CURR_SMOKE,  HYPERTENSION) %>%
    select(SEX, BP_TREAT, CURR_SMOKE, HYPERTENSION) %>%
    summarise(n = n()) %>%
    arrange(desc(n)) %>%
    kable
```

@@

```{r lm model}
lm(MAP ~ diversities_shannon, data=dset.alpha) %>%
    summary(.) %>% coef %>% kable(., digits=4)
lm(MAP ~ BL_AGE + SEX + diversities_shannon, data=dset.alpha) %>%
    summary(.) %>% coef %>% kable(., digits=4)
lm(MAP ~ BL_AGE + SEX + BMI + CURR_SMOKE + Q57X + BP_TREAT + diversities_shannon, data=dset.alpha) %>%
    summary(.) %>% coef %>% kable(., digits=4)

@@

Aaron work-december-2018 prepare_maaslin_input.R

@@

```{r scp html}
system('scp rrbiome.html utu:articletwo')
```

>>>>

filter.phenotype.data <- function(pseq,
                                  included = c("DNA_OK", "MEN", "BL_AGE", "SYSTM", "DIASM", "BP_TREAT", "BMI",
                                               "CURR_SMOKE", "DIAB", "BL_USE_RX_C09", "Q57X", "BL_USE_RX_C03",
                                               "BL_USE_RX_C07", "BL_USE_RX_C08"))
    meta(pseq) %>%
        tibble::rownames_to_column(var = "rowname") %>%
        dplyr::select(rowname, included) %>%
        dplyr::mutate(PULSEPRESSURE = SYSTM - DIASM,
                      HYPERTENSION = factor(ifelse(SYSTM >= 140 | DIASM >= 90 | BP_TREAT == 1, 1, 0)),
                      SEX = factor(ifelse(MEN == "Female", 1, 0)),
                      MAP = 2./3.*DIASM + 1./3.*SYSTM) %>%
 #       dplyr::filter(DNA_OK == 1) %>%
        stats::na.omit() %>%
        dplyr::select(-MEN, -DNA_OK, -BP_TREAT) %>%
        tibble::remove_rownames() %>%
        tibble::column_to_rownames(var = "rowname")
```

@@

Number of participants in each combination of all categorical variables

```{r numbers by group}
dset.species %>%
    group_by(SEX, CURR_SMOKE,  HYPERTENSION, Q57X,
             BL_USE_RX_C09, BL_USE_RX_C03, BL_USE_RX_C07, BL_USE_RX_C08) %>%
    select(-SYSTM, -DIASM, -BL_AGE, -BMI, -PULSEPRESSURE, -MAP) %>%
    summarise(n = n()) %>%
    arrange(desc(n)) %>%
    kable            
```

Number of particiapants in each combination after combining RR
medication and exercice to binomial variables.

```{r numbers by group reduced}
dset.species %>%
    mutate(ANYDRUG = ifelse(BL_USE_RX_C09  == 1 | BL_USE_RX_C03 == 1 |
                            BL_USE_RX_C07 == 1  | BL_USE_RX_C08 == 1, 1, 0),
           EXERCICE = ifelse(Q57X == 1, 0, 1)) %>%
    group_by(SEX, CURR_SMOKE,  HYPERTENSION, ANYDRUG, EXERCICE) %>%
    select(SEX, CURR_SMOKE, HYPERTENSION, EXERCICE) %>%
    summarise(n = n()) %>%
    arrange(desc(n)) %>%
    kable

dset.species %>% pull(Q57X) %>% summary
```

@@

```{r lm model}
lm(MAP ~ diversities_shannon, data=dset.alpha) %>%
    summary(.) %>% coef %>% kable(., digits=4)
lm(MAP ~ BL_AGE + SEX + diversities_shannon, data=dset.alpha) %>%
    summary(.) %>% coef %>% kable(., digits=4)
lm(MAP ~ BL_AGE + SEX + BMI + CURR_SMOKE + Q57X + diversities_shannon, data=dset.alpha) %>%
    summary(.) %>% coef %>% kable(., digits=4)

@@

Aaron work-december-2018 *prepare_maaslin_input.R*

@@

```{r scp html, cache = FALSE}
system('scp rrbiome.html utu:articletwo')
```
